A Bayesian Perspective \\\large{Extracorporeal CPR for Refractory Out-of-Hospital Cardiac Arrest}

### 2023-02-11 manuscript/manuscript.qmd
updated discussion

This Bayesian analysis of the INCEPTION ITT data alone (with a vague prior) suggests the presence of a non-negligible probability of benefit with eCPR, in the range of 73%, compared to cCPR. The probability that this improvement in survival exceeds 10% is almost 66%. The consideration of pertinent prior information from two previous RCTs[@RN6751;@RN6759] in the field results in estimates of improved eCPR survival probabilities, with intact neurological status, in the vicinity of 80-85%. Moreover depending on the choice of prior beliefs included, the posterior probability of at least a 10% improved eCPR survival is in the range of 71 - 77% probability. In contrast to the INCEPTION original conclusion of similar survival effects between the two treatment, this re-analysis suggests only a modest probability of approximately that survival probabilities between the two techniques are within a reasonably acceptable equivalence range. The overall message would appear to be that while any definitive conclusions regarding the superiority of eCPR are lacking, the possibility of a clinically meaningful benefit or even less likley clinically meaningful harm has not been reasonably excluded and continued research is necessary to clarifying the remaining uncertainties.

The INCEPTION researchers are to be congratulated on their trial design, its execution, and addressing an important clinical question in the most challenging of research environments. However the natural constraints of standard statistical analyses, also known as null hypothesis significance testing, limits the full exploitation of the current data and updating of past knowledge. For trails failing to meet statistical significance, those incorrectly labelled "negative" trials, null hypothesis significance testing favors confusion "between absence of evidence and evidence of absence"[@RN6765].

>>>>

This Bayesian analysis of the INCEPTION ITT data alone (with a vague prior) suggests the presence of a non-negligible probability of benefit with eCPR, in the range of 73%, compared to cCPR. The probability that this improvement in survival exceeds 10% is almost 66%. The consideration of pertinent prior information from two previous RCTs[@RN6751;@RN6759] in the field results in estimates of improved eCPR survival probabilities, with intact neurological status, in the vicinity of 80-85%. Moreover depending on the choice of prior beliefs included, the posterior probability of at least a 10% improved eCPR survival is in the range of 71 - 77% probability. In contrast to the INCEPTION original conclusion of similar survival effects between the two treatments, this re-analysis suggests only a modest probability of approximately 15% that survival probabilities between the two techniques are within a reasonably acceptable equivalency range. Additional uncertainty about the benefits or harms of eCPR arise from the per protocol analysis, both with and without informative priors.

The INCEPTION researchers are to be congratulated on their trial design, its execution, and addressing an important clinical question in the most challenging of research environments. However the natural constraints of standard statistical analyses, also known as null hypothesis significance testing, limits the full exploitation of the current data and updating of past knowledge. For trials such as INCEPTION that fail to meet statistical significance and are often thought of as "negative" trials, null hypothesis significance testing favors confusion "between absence of evidence and evidence of absence"[@RN6765]. Rather than dichotomizing results into statistical significance or not, with its obligatory loss of information, Bayesian analysis concentrates on estimation of key outcome differences with direct probability measures of their uncertainty.

Although this  Bayesian analysis presents the strengths noted above, it also had limitations. First, we did not have access to either
individual data from any of the trials limiting the possibility of examining specific subgroups. Also for the per protocol analysis, this data was only available for survival and not the primary outcome that included survival with intact neurological status. The threshold choices for superiority and equivalence may be seen as arbitrary, although the analyses can be easily repeated for different choices. Given the existence of only 3 trials  differences between the baseline cCPR rates could not be reliably modeled. We choose not to perform a traditional Bayesian meta-analysis with
a semi-informative prior for the between-study variation  because our
goal was to illustrate how Bayesian principles could be informative
when applied to the interpretation of new trial data.

The overall message from this study is that standard statistical analyses leading to a conclusion of "similar survival effects" is overly simplified and potentially inaccurate. This Bayesian analysis demonstrates clearly that while any definitive conclusions regarding the superiority of eCPR are lacking, the possibility of a clinically meaningful benefit, or even somewhat less likely clinically meaningful harm, has not been reasonably excluded and continued research is necessary to clarifying the remaining uncertainties.

### 2023-02-11 manuscript/manuscript.tex
updated discussion

\begin{keyword}
    extracorporeal CPR \sep 
    Bayesian statistics
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, sharp corners, enhanced, interior hidden, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, frame hidden]}{\end{tcolorbox}}\fi

@@

This Bayesian analysis of the INCEPTION ITT data alone (with a vague
prior) suggests the presence of a non-negligible probability of benefit
with eCPR, in the range of 73\%, compared to cCPR. The probability that
this improvement in survival exceeds 10\% is almost 66\%. The
consideration of pertinent prior information from two previous
RCTs\citep{RN6751, RN6759} in the field results in estimates of improved
eCPR survival probabilities, with intact neurological status, in the
vicinity of 80-85\%. Moreover depending on the choice of prior beliefs
included, the posterior probability of at least a 10\% improved eCPR
survival is in the range of 71 - 77\% probability. In contrast to the
INCEPTION original conclusion of similar survival effects between the
two treatment, this re-analysis suggests only a modest probability of
approximately that survival probabilities between the two techniques are
within a reasonably acceptable equivalence range. The overall message
would appear to be that while any definitive conclusions regarding the
superiority of eCPR are lacking, the possibility of a clinically
meaningful benefit or even less likley clinically meaningful harm has
not been reasonably excluded and continued research is necessary to
clarifying the remaining uncertainties.

The INCEPTION researchers are to be congratulated on their trial design,
its execution, and addressing an important clinical question in the most
challenging of research environments. However the natural constraints of
standard statistical analyses, also known as null hypothesis
significance testing, limits the full exploitation of the current data
and updating of past knowledge. For trails failing to meet statistical
significance, those incorrectly labelled ``negative'' trials, null
hypothesis significance testing favors confusion ``between absence of
evidence and evidence of absence''\citep{RN6765}.

>>>>

\begin{keyword}
    extracorporeal CPR \sep 
    Bayesian statistics
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, sharp corners, breakable, interior hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, boxrule=0pt]}{\end{tcolorbox}}\fi

@@

This Bayesian analysis of the INCEPTION ITT data alone (with a vague
prior) suggests the presence of a non-negligible probability of benefit
with eCPR, in the range of 73\%, compared to cCPR. The probability that
this improvement in survival exceeds 10\% is almost 66\%. The
consideration of pertinent prior information from two previous
RCTs\citep{RN6751, RN6759} in the field results in estimates of improved
eCPR survival probabilities, with intact neurological status, in the
vicinity of 80-85\%. Moreover depending on the choice of prior beliefs
included, the posterior probability of at least a 10\% improved eCPR
survival is in the range of 71 - 77\% probability. In contrast to the
INCEPTION original conclusion of similar survival effects between the
two treatments, this re-analysis suggests only a modest probability of
approximately 15\% that survival probabilities between the two
techniques are within a reasonably acceptable equivalency range.
Additional uncertainty about the benefits or harms of eCPR arise from
the per protocol analysis, both with and without informative priors.

The INCEPTION researchers are to be congratulated on their trial design,
its execution, and addressing an important clinical question in the most
challenging of research environments. However the natural constraints of
standard statistical analyses, also known as null hypothesis
significance testing, limits the full exploitation of the current data
and updating of past knowledge. For trials such as INCEPTION that fail
to meet statistical significance and are often thought of as
``negative'' trials, null hypothesis significance testing favors
confusion ``between absence of evidence and evidence of
absence''\citep{RN6765}. Rather than dichotomizing results into
statistical significance or not, with its obligatory loss of
information, Bayesian analysis concentrates on estimation of key outcome
differences with direct probability measures of their uncertainty.

Although this Bayesian analysis presents the strengths noted above, it
also had limitations. First, we did not have access to either individual
data from any of the trials limiting the possibility of examining
specific subgroups. Also for the per protocol analysis, this data was
only available for survival and not the primary outcome that included
survival with intact neurological status. The threshold choices for
superiority and equivalence may be seen as arbitrary, although the
analyses can be easily repeated for different choices. Given the
existence of only 3 trials differences between the baseline cCPR rates
could not be reliably modeled. We choose not to perform a traditional
Bayesian meta-analysis with a semi-informative prior for the
between-study variation because our goal was to illustrate how Bayesian
principles could be informative when applied to the interpretation of
new trial data.

The overall message from this study is that standard statistical
analyses leading to a conclusion of ``similar survival effects'' is
overly simplified and potentially inaccurate. This Bayesian analysis
demonstrates clearly that while any definitive conclusions regarding the
superiority of eCPR are lacking, the possibility of a clinically
meaningful benefit, or even somewhat less likely clinically meaningful
harm, has not been reasonably excluded and continued research is
necessary to clarifying the remaining uncertainties.
